---
title: "Running Analysis"
output: html_document
---

```{r setup, include=FALSE}
library(lubridate)
library(tidyverse)
library(skimr)
library(leaflet)
source("Code/0_Import_Data.R")
```

```{r}
summary_data <- summary(all_runs)
```


When do I start running?

```{r}
summary_data <- summary_data %>%
  mutate(
    start_time = hms(format(sessionStart, format = "%H:%M:%S")),
    run_date = ymd(format(sessionStart, format = "%Y-%m-%d")),
    start_time_frac = hour(start_time) + minute(start_time) / 60
  )

summary_data %>%
  ggplot(aes(x = run_date, y = start_time_frac)) +
  geom_smooth(method = "loess") +
  geom_point(size = 1) +
  theme_classic(base_size = 16) +
  scale_y_continuous(breaks = 0:18) +
  labs(
    title = "Starting times for runs",
    x = "Date", y = "Starting time"
  )
```
There is a run at 2am, which doesn't seem right. What is going on there?

```{r}
summary_data %>% filter(start_time_frac < 3)
```



Run duration over time

```{r}
summary_data %>%
  ggplot(aes(x = run_date, y = duration)) +
  geom_smooth(method = "loess") +
  geom_point(size = 1) +
  theme_classic(base_size = 16) +
  labs(
    title = "Run Duration",
    x = "Date", y = "Run Duration (mins)"
  )
```

Run distance over time

```{r}
summary_data %>%
  ggplot(aes(x = run_date, y = distance / 1000)) +
  geom_smooth(method = "loess") +
  geom_point(size = 1) +
  theme_classic(base_size = 16) +
  labs(
    title = "Run Distance",
    x = "Date", y = "Run Distance (km)"
  )
```

Average speed over time 

```{r}
summary_data <- summary_data %>%
  mutate(average_speed = (distance / 1000) / 
           (as.numeric(duration) / 60))

summary_data %>%
  ggplot(aes(x = run_date, y = average_speed)) +
  geom_smooth(method = "loess") +
  geom_point(size = 1) +
  theme_classic(base_size = 16) +
  labs(
    title = "Average Speed",
    x = "Date", y = "Average Speed (km/h)"
  )
```

Average pace over time 

```{r}
summary_data %>%
  ggplot(aes(x = run_date, y = avgPaceMoving)) +
  geom_smooth(method = "loess") +
  geom_point(size = 1) +
  theme_classic(base_size = 16) +
  labs(
    title = "Average Pace",
    x = "Date", y = "Average Pace (min/km)"
  )
```

Run Pace vs distance

```{r}
summary_data %>%
  ggplot(aes(x = distance / 1000, y = avgPaceMoving)) +
  geom_smooth(method = "loess") +
  geom_point(size = 1) +
  theme_classic(base_size = 16) +
  labs(
    title = "Run Pace vs distance",
    y = "Run Pace (min/km)", x = "Distance (km)"
  )
```

Pace vs elevation gain

```{r}
summary_data %>%
  ggplot(aes(x = total_elevation_gain, y = avgPaceMoving)) +
  geom_smooth(method = "loess") +
  geom_point(size = 1) +
  theme_classic(base_size = 16) +
  labs(
    title = "Run Pace vs Elevation Gain",
    y = "Run Pace (min/km)", x = "Elevation Gain (m)"
  )
```

How hot is it when I run?

```{r}
summary_data %>% 
  ggplot(aes(x = avgHeartRate)) + geom_histogram() +
  theme_classic(base_size = 16) +
  labs(
    title = "Average Heart Rate when running",
    y = "Frequency", x = "Average Heart Rate (bpm)"
  )
```
Average Heart Rate over time 

```{r}
summary_data %>%
  filter(run_date > ymd('2019-10-01')) %>% 
  ggplot(aes(x = run_date, y = avgHeartRate)) +
  geom_smooth(method = "loess") +
  geom_point(size = 1) +
  theme_classic(base_size = 16) +
  labs(
    title = "Average Heart Rate",
    x = "Date", y = "Average Heart Rate (bpm)"
  )
```

Run Frequency by month

```{r}
monthly_summary <- summary_data %>% 
  mutate(year_month = ym(format(run_date, format = "%Y-%m"))) %>% 
  group_by(year_month) %>% 
  summarise(
    distance = sum(distance, na.rm = T),
    duration = sum(duration, na.rm = T),
    pace = mean(avgPaceMoving),
    wrRatio = mean(wrRatio),
    Num_Run = n()
  )


monthly_summary %>% 
  ggplot(aes(x = year_month, y = Num_Run)) +
  geom_bar(stat = 'identity', colour = 'blue') +
  theme_classic(base_size = 16) +
  labs(
    title = "Number of runs per month",
    x = "Month", y = "Number of runs"
  )

monthly_summary %>% 
  ggplot(aes(x = year_month, y = distance / 1000)) + 
  geom_line(stat = 'identity', colour = 'blue') +
  theme_classic(base_size = 16) +
  labs(
    title = "Distance run per month",
    x = "Month", y = "Distance Run (km)"
  )

monthly_summary %>% 
  ggplot(aes(x = year_month, y = duration / 60)) + 
  geom_line(stat = 'identity', colour = 'blue') +
  theme_classic(base_size = 16) +
  labs(
    title = "Time spent running per month",
    x = "Month", y = "Time spent running (hr)"
  )

monthly_summary %>% 
  ggplot(aes(x = year_month, y = pace)) + 
  geom_line(stat = 'identity', colour = 'blue') +
  theme_classic(base_size = 16) +
  labs(
    title = "Average pace per month",
    x = "Month", y = "Average pace (min/km)"
  )

monthly_summary %>% 
  ggplot(aes(x = year_month, y = wrRatio)) + 
  geom_line(stat = 'identity', colour = 'blue') +
  theme_classic(base_size = 16) +
  labs(
    title = "Work to rest ratio per month",
    x = "Month", y = "Work to rest ratio"
  )
```


Smoothed run frequency

Break down performance by year


```{r}
skim(summary_data)
```

# Look at GPS Data


```{r}
plot(all_runs, session = 482)
plotRoute(all_runs, session = 481:482, zoom = 15)
# leaflet
```

```{r}
m <- leaflet() %>%
  addTiles()

m
```

```{r}
head(all_runs[[482]],n=30)
```

```{r}
leaflet() %>% 
  addTiles()
```


Plan of analysis
* Can we identify similar routes? There are some fairly similar/standard runs
 * If so, can we look at change in performance over time
* Can I identify the fastest run of a certain distance
  * Is there a typical split pattern for distances?
* Can I identify the fastest run of a certain route
* Can I identify where I start too fast?
* Can I make my own global heatmap?
